---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
```


```{r}

data_path <- "C:/Users/Admin/Desktop/STATS/"

setwd(data_path)

# load the necessary packages ("extensions")
library(tidyverse)
library(lubridate) # to work with dates
library(skimr) # for summaries of data
library(arrow) # to be able to load data in the .parquet format
library(gender)
# read application data

app_data_sample <- read_parquet(paste0(data_path,"app_data_sample.parquet"))

app_data_sample

```

```{r}
examiner_names <- app_data_sample %>% 
  distinct(examiner_name_first)
examiner_names

```

```{r gender-2}
# get a table of names and gender
library(gender)
library(genderdata)
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )
examiner_names_gender

```

```{r gender-3}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)
# joining gender back to the dataset
app_data_sample <- app_data_sample %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")
# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```

```{r race-1}
library(wru)
examiner_surnames <- app_data_sample %>% 
  select(surname = examiner_name_last) %>% 
  distinct()
examiner_surnames
```

```{r race-2}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
examiner_race
```

```{r race-3}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))
examiner_race
```

```{r race-4}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)
app_data_sample <- app_data_sample %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))
rm(examiner_race)
rm(examiner_surnames)
gc()
```

```{r tenure-1}
library(lubridate) # to work with dates
examiner_dates <- app_data_sample %>% 
  select(examiner_id, filing_date, appl_status_date) 
examiner_dates
```


```{r tenure-2}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date))) %>% 
  filter(year(end_date)<2018)
```



```{r tenure-3}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    )
examiner_dates
```



```{r tenure-4}
app_data_sample <- app_data_sample %>% 
  left_join(examiner_dates, by = "examiner_id")
rm(examiner_dates)
gc()
```



```{r sum-1}
subset_app_data <- app_data_sample %>% 
  mutate(race = as_factor(race), gender = as_factor(gender)) %>% 
  select(gender, race, tenure_days) 
subset_app_data %>% 
  count(gender) %>% 
  mutate(pct = n/sum(n))
```

```{r sum-2}
subset_app_data %>% 
  count(race) %>% 
  mutate(pct = n/sum(n))
```



```{r sum-3}
library(skimr)
subset_app_data %>%  
  skim()
```



```{r plot-gender-1, warning=FALSE}
person_level_data <- app_data_sample %>% 
  group_by(examiner_id) %>% 
  summarise(
    art_unit = min(examiner_art_unit, na.rm = TRUE),
    gender = min(gender, na.rm = TRUE),
    start_year = min(year(earliest_date), na.rm = TRUE),
    latest_date = max(latest_date, na.rm = TRUE),
    tenure_days = max(tenure_days, na.rm = TRUE)
  ) %>% 
  mutate(
    tc = floor(art_unit/100)*100,
    work_group = floor(art_unit/10)*10
  ) %>% 
  filter(!is.na(gender)) # dropping all records where we don't know the gender
person_level_data
```



```{r plot-gender-2}
ggplot(person_level_data) +
  geom_boxplot(aes(x = tenure_days, color = gender))
```



```{r plot-gender-3}
ggplot(person_level_data) +
  geom_bar(
    aes(x=as_factor(tc), fill = gender), 
    position = position_stack()
    ) +
  xlab("Technology Center")
```


```{r plot-gender-4}
ggplot(person_level_data) +
  geom_bar(
    aes(x=as_factor(work_group), fill = gender), 
    position = position_stack()
    ) +
  xlab("Work group") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) + #rotate labels
  facet_wrap(vars(tc), scales = "free")
```

```{r}
library(dplyr)
library(ggplot2)

lm.breakdown=lm(tenure_days ~ gender+race, data=subset_app_data)
summary(lm.breakdown)

```

```{r}
library(dplyr)
library(ggplot2)

person_level_data$tc=as.factor(person_level_data$tc)
lm.breakdown2=lm(tenure_days ~ tc+art_unit, data=person_level_data)
summary(lm.breakdown2)

```


Based on the Regression analysis, it appears being Hispanic has the strongest negative correlation with Tenure, while being Black has the most positive correlation with tenure. Being Male has a negative correlation as well.

However, none of these correlations are too significant, with the highest absolute value being +394.

The correct way to do this would have been to assign Male and Female a value of 1 and 0 respectively, and Races a number 1 through 5 as well, and have the model run through it. I do not know if the model does it automatically or no but have assumed so for this exercise.

For the second Model, i used the TC and Art_unit as variables. The model seems to be wrong because the values seem really high, unless there really is a positive correlation between the number of years worked per TC. TC 2400 seems to be where people work the longest.

Not sure what kind of recommendation to make as race seems to be insignificant, but maybe look into what makes TC2400 so popular where people chose to stay longer and seek to replicate it in other TCs.



