---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r prep-steps}
# set path for R to find our data
data_path <- "C:/Users/Admin/Desktop/STATS/"
library(arrow) # to be able to load data in the .parquet format
# read application data
app_data_sample <- read_parquet(paste0(data_path,"app_data_sample.parquet"))
library(gender)
#install_genderdata_package() # only run this line the first time you use the package, to get data for it
# get a list of first names without repetitions
examiner_names <- app_data_sample %>% 
  distinct(examiner_name_first)
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)
# joining gender back to the dataset
app_data_sample <- app_data_sample %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")
# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
# Examiners' race
library(wru)
examiner_surnames <- app_data_sample %>% 
  select(surname = examiner_name_last) %>% 
  distinct()
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))
examiner_race <- examiner_race %>% 
  select(surname,race)
app_data_sample <- app_data_sample %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))
rm(examiner_race)
rm(examiner_surnames)
gc()
# Examiner's tenure
library(lubridate) # to work with dates
examiner_dates <- app_data_sample %>% 
  select(examiner_id, filing_date, appl_status_date) 
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date))) %>% 
  filter(year(end_date)<2018) %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    )
app_data_sample <- app_data_sample %>% 
  left_join(examiner_dates, by = "examiner_id")
rm(examiner_dates)
gc()
```
## Adding paygrade data

First, we load the paygrade file.

```{r load-paygrade-data}
examiner_gs <- read_csv(paste0(data_path,"examiner_gs.csv"))
examiner_ids <- read_csv(paste0(data_path,"examiner_ids.csv"))
```

```{r join-id}
examiner_gs <- examiner_gs %>% 
  left_join(examiner_ids) %>% 
  select(
    grade = examiner_grade,
    start_date,
    end_date,
    examiner_id = patex_id
  )
```

```{r get-time-in-grade}
time_in_grade <- examiner_gs %>% 
  mutate(
    start_date = mdy(start_date), # converting into proper date type
    end_date = mdy(end_date), # converting into proper date type
    days_in_grade = interval(start_date, end_date) %/% days(1)
  ) %>% 
  group_by(examiner_id) %>% 
  filter(grade!=max(grade, na.rm = TRUE)) %>% # dropping the highest grade record
  summarise(mean_days_in_grade = mean(days_in_grade, na.rm = TRUE))
time_in_grade
```

```{r app-proc-time}
examiner_data <- app_data_sample %>% 
  filter(disposal_type!="PEND") %>% # here, we exclude in-process applications
  mutate(
    app_start_date = ymd(filing_date),
    app_end_date = case_when(
      disposal_type == "ISS" ~ ymd(patent_issue_date), # for issued patents
      disposal_type == "ABN" ~ ymd(abandon_date), # for abandoned applications
      TRUE ~ NA_Date_
    ),
    app_proc_days = interval(app_start_date, app_end_date) %/% days(1)) %>% 
  filter(app_proc_days>0 & app_proc_days < 3650) %>% # limit to 0-10 years
  group_by(examiner_id) %>% 
  summarise(
    app_count = n(),
    tc = min(tc, na.rm = TRUE),
    gender = first(gender),
    race = first(race),
    tenure_days = max(tenure_days, na.rm = TRUE),
    mean_app_proc_days = mean(app_proc_days, na.rm = TRUE)
  )
examiner_data
```

```{r add-time-in-grade}
examiner_data <- examiner_data %>% 
  left_join(time_in_grade)
examiner_data
```

```{r reg-sample}
library(modelsummary)
models <- list()
models[['m1']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days, data = examiner_data) 
models[['m2']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days + as_factor(race), 
    data = examiner_data) 
models[['m3']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days + as_factor(gender), 
         data = examiner_data) 
modelsummary(models)
```

```{r gender-check}
women_variable <- examiner_data %>% 
filter(gender == "female")
mean(women_variable$mean_days_in_grade,na.rm=TRUE)
men_variable <- examiner_data %>% 
filter(gender == "male")
mean(men_variable$mean_days_in_grade,na.rm=TRUE)
```

```{r race-check}
white_variable <- examiner_data %>% 
  filter(race == "white")
mean(white_variable$mean_days_in_grade,na.rm=TRUE)
asian_variable <- examiner_data %>% 
  filter(race == "Asian")
mean(asian_variable$mean_days_in_grade,na.rm=TRUE)
black_variable <- examiner_data %>% 
  filter(race == "black")
mean(black_variable$mean_days_in_grade,na.rm=TRUE)
hispanic_variable <- examiner_data %>% 
  filter(race == "Hispanic")
mean(hispanic_variable$mean_days_in_grade,na.rm=TRUE)
```


There seems to be no real difference for Gender when it comes to promotion.
When it comes to race They also seem closely related, however black people tend to take longest at 589 days and hispanic the least at 504 days.

From the means and regression model summary, there seems to not be any effect of gender on race on the time it takes to get a promotion.

There could be some limitations such as:
Not standardized method of promoting people. Different examiners or promoters could have different biases.
Other Factors such as work ethic/production that are better indicators of promotion.
Assumes they all do the same job or department. Different departments could have different criteria or dominant race/gender combos.