---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)
library(tidygraph)
library(ggraph)

# define nodes
node_names <- tibble(
  id   = c(1,2,3,4,5,6,7,8,9,10),
  name = c("1","2","A","B","C","6","D","3","4","5")
)
node_names
```


```{r}
# define connections (have to correspond to ties 1-2, 2-3, 2-4, 3-4)
# for each element in `from` there is a corresponding element in `to`
edge_list <- tibble(
  from = c(1,2,3,3,4,4,4,4,5,5,5,6,6,7,7,8,8),
  to   = c(2,3,4,5,5,6,7,8,7,8,9,7,10,8,10,9,10)
)
edge_list
```

```{r}
# combine this information into a network graph object
friendship_graph <- tbl_graph(nodes = node_names, edges = edge_list, directed = FALSE)
friendship_graph
```

```{r}
friendship_graph %>% 
    ggraph(layout = 'kk') + 
    geom_edge_link() + 
    geom_node_point(size = 8, colour = 'gray') +
    geom_node_text(aes(label = name), colour = 'steelblue', vjust = 0.4) + 
    ggtitle('Friendship network') + 
    theme_graph()
```

```{r}
friendship_graph <- friendship_graph %>% 
  activate(nodes) %>% # we need to state we'll be adding to nodes, not edges
  mutate(d_centrality = centrality_degree()) %>%  # adding measure of degree centrality
  mutate(b_centrality = centrality_betweenness()) # adding betweenness centrality
## Warning in betweenness(graph = graph, v = V(graph), directed = directed, :
## 'nobigint' is deprecated since igraph 1.3 and will be removed in igraph 1.4
friendship_graph
```

```{r}
friendship_graph %>% 
  ggraph(layout = 'kk') + 
  geom_edge_link() + 
  geom_node_point(aes(size = d_centrality, colour = b_centrality)) + 
  scale_color_continuous(guide = 'legend') +
  geom_node_text(aes(label = name), colour = 'red', vjust = 1.6) + 
  ggtitle('Friendship network') + 
  theme_graph()
```

Based on the Colours shown on the graph, A seems to be the most central. It acts as a broker between the rest of the network and 1 and 2. A is also connected to B and C which are the 2 biggest nodes with the most connections at 5 each. A can reach anybody in the network within 3 jumps.

The next best options are B and C in terms of the nu,bers of connections they have as well as their appeared centrality within the network, having a large number of connections.

